Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-01-23 ICML-submission/mistake-lower-bound-for-ignorant-algorithms.tex
.

In this section, we prove a mistake lower bound for a family of algorithms
called \textit{ignorant algorithms}. Ignorant algorithms ignore the examples on
which they make mistakes. This assumption seems strong, but as we will explain
below, it is actually natural, and several recently proposed bandit
classification algorithms that achieve the $\sqrt{T}$-regret bound belong to
this family, e.g., SOBA~\citep{Beygelzimer-Orabona-Zhang-2017},
OBAMA~\citep{Foster-Kale-Luo-Mohri-Sridharan-2018}. Also,
\textsc{Nearest-Neighbor Algorithm} (Algorithm~\ref{algorithm:nearest-neighbor})
presented in Appendix~\ref{section:nearest-neighbor-algorithm} is an ignorant
algorithm.

@@

The ignorant algorithm is formally defined as follows.
\begin{definition}
\label{definition:ignorant-algorithm}
An algorithm $\calA$ for \textsc{Online Multiclass Linear Classification with
Bandit Feedback} is called \emph{ignorant} if for all $t \in \{1,2,\dots,T\}$,
$\widehat y_t$ is drawn from some distribution that is jointly determined by
\begin{enumerate}
\item $x_t$
\item $H_t \triangleq \left( (x_{\tau_1}, y_{\tau_1}), (x_{\tau_2}, y_{\tau_2}), \ldots, (x_{\tau_{n_t}}, y_{\tau_{n_t}}) \right)$,
where $1 \le \tau_1 < \dots < \tau_{n_t} < t$, and $\{\tau_1, \ldots, \tau_{n_t}\}$
are the rounds in which the algorithm predicts correctly (that is,
$\{\tau_1, \dots, \tau_{n_t}\} = \{1 \le \tau < t ~:~ \widehat y_\tau=y_\tau\}$).
\end{enumerate}
We use $p_t(\cdot|x_t)\in \{p\in \R_+^K ~:~ p^\top\mathbf{1}=1\}$ to denote
the distribution $\widehat y_t$ is drawn from. Note that we let the current
feature vector $x_t$ be an argument to $p_t$ to incorporate its dependency on
$x_t$. Therefore, for an ignorant algorithm, the function $p_t(\cdot|\cdot)$ is
only determined by $H_t$.
\end{definition}

@@

\begin{theorem}[Mistake Lower Bound for Ignorant Algorithms]
\label{theorem:ignorant-lower-bound}
Suppose $\calA$ is an ignorant algorithm for \textsc{Online Multiclass Linear
Classification with Bandit Feedback}. There exists an adversary that can
sequentially choose labeled examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)
\in \R^d\times \{1,2\}$ so that the examples are $\gamma$-strongly linearly
separable, $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$, and the expected
number of mistakes made by $\calA$ is at least
$$
\frac{1}{10}\min\left\{\sqrt{T}, \left(\frac{1}{160\gamma}\right)^{\frac{d-2}{4}} \right\} \; .
$$
\end{theorem}

>>>>

In this section, we prove a mistake lower bound for a family of algorithms
called \textit{ignorant algorithms}. Ignorant algorithms ignore the examples on
which they make mistakes. This assumption seems strong, but as we will explain
below, it is actually natural, and several recently proposed bandit
classification algorithms that achieve $\sqrt{T}$ regret bounds belong to
this family, e.g., SOBA~\citep{Beygelzimer-Orabona-Zhang-2017},
OBAMA~\citep{Foster-Kale-Luo-Mohri-Sridharan-2018}. Also,
\textsc{Nearest-Neighbor Algorithm} (Algorithm~\ref{algorithm:nearest-neighbor})
presented in Appendix~\ref{section:nearest-neighbor-algorithm} is an ignorant
algorithm.

@@

The ignorant algorithm is formally defined as follows.
\begin{definition}
\label{definition:ignorant-algorithm}
An algorithm $\calA$ for \textsc{Online Multiclass Linear Classification with
Bandit Feedback} is called \emph{ignorant} if for all $t \in \{1,2,\dots,T\}$,
$\widehat y_t$ is drawn from some distribution that is jointly determined by
\begin{enumerate}
\item $x_t$,
\item $H_t \triangleq \left( (x_{\tau_1}, y_{\tau_1}), (x_{\tau_2}, y_{\tau_2}), \ldots, (x_{\tau_{n_t}}, y_{\tau_{n_t}}) \right)$,
where $1 \le \tau_1 < \dots < \tau_{n_t} < t$, and $\{\tau_1, \ldots, \tau_{n_t}\}$
are the rounds in which the algorithm predicts correctly (that is,
$\{\tau_1, \dots, \tau_{n_t}\} = \{1 \le \tau < t ~:~ \widehat y_\tau=y_\tau\}$).
\end{enumerate}
We use $p_t(\cdot|x_t)\in \{p\in \R_+^K ~:~ p^\top\mathbf{1}=1\}$ to denote
the distribution $\widehat y_t$ is drawn from. Note that we let the current
feature vector $x_t$ be an argument to $p_t$ to incorporate its dependency on
$x_t$. Therefore, for an ignorant algorithm, the function $p_t(\cdot|\cdot)$ is
only determined by $H_t$.
\end{definition}

@@

\begin{theorem}[Mistake lower bound for ignorant algorithms]
\label{theorem:ignorant-lower-bound}
Suppose $\calA$ is an ignorant algorithm for \textsc{Online Multiclass Linear
Classification with Bandit Feedback}. There exists an adversary that can
sequentially choose labeled examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)
\in \R^d\times \{1,2\}$ so that the examples are $\gamma$-strongly linearly
separable, $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$, and the expected
number of mistakes made by $\calA$ is at least
$$
\frac{1}{10}\min\left\{\sqrt{T}, \left(\frac{1}{160\gamma}\right)^{\frac{d-2}{4}} \right\} \; .
$$
\end{theorem}
